Tabular Constraint Learning

### 2016-04-12 Paper/ECAI/tabular.tex
intro sketch

\begin{abstract}
  abstract
\end{abstract}
\section{Introduction}
\sergey{bullet points for luc to start introduction}\\
\textbf{Key question}:\\
Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\

\textbf{Motivation}:
\samuel{Suggest better table layouts / formula translation}
\begin{itemize}
  \item File generated from model, model got lost, need to reconstruct
  \item Constraint programming is hard - is Excel hard?
  \item Avoid manual analysis, provide selection of constraints
  \item Error checking
  \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
\end{itemize}

>>>>

\begin{abstract}
  abstract
\end{abstract}
\section{Introduction}
Imagine you have received a CSV file. Upon examining the content it becomes fairly clear it was made in Excel. Even the names of the fields such as Total or Avg indicated the usage of the standard functions. Having a closer look, you could deduce the functional dependencies in the table that look as in Figure \ref{fig:main_example}. Provided that a spreadsheet and the numbers in it are small, and dependencies are not complicated, one can examine the spreadsheet and find all connections. But wouldn't it be good to do it automatically?

This brings us to the \textbf{key question}:\\
Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\

Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task. We know that often people work in Excel with ranges, e.g. if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns. Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2. I.e., all operations are vectorized and make use of the standard Excel functions.

It is not hard to see that computation goes not only in columns but in rows as well: if we sum the values in Table 1 in the first row in all four quarters, we would obtain the seventh column in Table 1. That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns. Here everything is mixed, some values are missing in the results, some missing the arguments, even a header might have missing values. The data is relational on one hand, since we have multiple table with relationships between them. And on the other hand, the data is numeric, since people tend to use spreadsheets for financial and accounting computations, which is heavily numeric.

\sergey{we need to incorporate the motivation part like error checking and autocompletion in the intro}

\sergey{bullet points for Luc and Tias to rework introduction}\\
\textbf{Motivation}:
\samuel{Suggest better table layouts / formula translation}
\begin{itemize}
  \item File generated from model, model got lost, need to reconstruct
  \item Constraint programming is hard - is Excel hard?
  \item Avoid manual analysis, provide selection of constraints
  \item Error checking
  \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
\end{itemize}
